---
title: "Air pollution and agricultural productivity in a developing country"
subtitle: 
author: 
  - "Joshua D. Merfeld <br> KDI School and IZA"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
                      fig.width = 9, fig.height = 3.5, fig.retina = 3,
                      out.width = "100%",
                      cache = FALSE,
                      echo = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      hiline = TRUE
                      )


library(sf)
library(sp)
library(tidyverse)
library(gganimate)
library(transformr)
library(gifski)
library(ggthemes)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(lubridate)
library(png)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)



# We want to get the wd up to the main folder
# Need to go up two levels
setwd('..')
# Double check
getwd()    # check


reportP <- function(pValue){
  if (pValue < 0.001){
    result <- "p < 0.001"
  } else {
    result <- sprintf("p = %.3f", pValue) # inserts a float into a string and simultaneously do rounding
  }
  return(result)
}

reportDec <- function(dec){
  
    result <- sprintf("%.3f", dec) # inserts a float into a string and simultaneously do rounding
  
  return(result)
}

reportDecOne <- function(dec){
  
    result <- sprintf("%.1f", dec) # inserts a float into a string and simultaneously do rounding
  
  return(result)
}

reportPercent <- function(dec){
  
    result <- sprintf("%.1f", 100*dec) # inserts a float into a string and simultaneously do rounding
  
  return(result)
}

commaNum <- function(large){
  
  
  result <- formatC(large, big.mark = ",")

  return(result)
}


# Shapefile of india
india_shape <- read_sf("data/spatial/shapefiles/state.shp")



# Coal plants
### Load raw data ----------------------------------------------------------------------------------------------------------------------------------------------------------
plants <- read.xlsx("data/raw/coal_plants.xlsx", sheet = "Units")
# India only
plants <- plants %>% filter(Country=="India")
# Also want only things with a non-missing year
plants <- plants %>% filter(is.na(Year)==F)

# Just keep what we want 
plants <- plants %>% dplyr::select(plant_id = ParentID, 
                                   unit_id = Tracker.ID, 
                                   capacity = `Capacity.(MW)`, 
                                   year_built = Year, 
                                   year_retired = RETIRED, 
                                   lat = Latitude, 
                                   lon = Longitude)


gps_points_plants <- plants %>% dplyr::select(lon, lat)
plants <- st_as_sf(SpatialPointsDataFrame(gps_points_plants, plants %>% dplyr::select(-c(lon, lat)), proj4string = CRS("EPSG:4326")))
plants <- plants %>% mutate(capacity = capacity/1000)

plants1990 <- plants %>% 
                  filter(year_built<=1990 & (year_retired>1990 | is.na(year_retired)==T)) %>%
                  group_by(plant_id) %>%
                  mutate(capacity = sum(capacity)) %>%
                  filter(row_number()==1) %>%
                  ungroup()
                  
plants2010 <- plants %>% 
                  filter(year_built<=2010 & (year_retired>2010 | is.na(year_retired)==T)) %>%
                  group_by(plant_id) %>%
                  mutate(capacity = sum(capacity)) %>%
                  filter(row_number()==1) %>%
                  ungroup()



# Wind example
# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
villages <- villages %>% filter(state=="23" & district=="07")
# Go through two months worth of days
date <- as_date("2010-01-01")
villageswind <- c()
for (add_day in 1:100){
  wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-", month(date), "-", day(date), ".csv")) %>% as_tibble()
  wind <- wind[,1:2]
  colnames(wind) <- c("shrid", "indicator")
  villagestemp <- villages %>% left_join(wind, by = "shrid")
  villagestemp$date <- date
  
  villageswind <- rbind(villageswind, villagestemp)
  
  date <- date + 1
}

villageswind <- villageswind %>% st_set_crs(st_crs(districts))



# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
village_shp <- villages
villages <- villages %>% filter(state=="23" & district=="07")
wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
wind$days_sum <- apply(wind[,2:8], 1, FUN = "sum")
wind <- wind %>% dplyr::select(shrid, days_sum)
villages <- villages %>% left_join(wind, by = "shrid")
villages <- villages %>% st_set_crs(st_crs(districts))
village_shp <- village_shp %>% st_set_crs(st_crs(districts))
# and crs for plants
plants1990 <- plants1990 %>% st_transform(st_crs(districts))
plants2010 <- plants2010 %>% st_transform(st_crs(districts))
india_shape2 <- india_shape %>% st_transform(st_crs(districts))

villages <- st_snap(villages, districts, tolerance = 1250)
crops_all <- read_csv(paste0("data/clean/ag_productivity/crop_area.csv"))

# plants

plants <- read.xlsx("data/raw/coal_plants.xlsx", sheet = "Units")
# India only
plants <- plants %>% filter(Country=="India")
# Also want only things with a non-missing year
plants <- plants %>% filter(is.na(Year)==F)

# Just keep what we want 
plants <- plants %>% dplyr::select(plant_id = ParentID, 
                                   unit_id = Tracker.ID, 
                                   capacity = `Capacity.(MW)`, 
                                   year_built = Year, 
                                   year_retired = RETIRED, 
                                   lat = Latitude, 
                                   lon = Longitude)

plants <- plants %>% mutate(capacity = capacity/1000)
# long
plantsyearly <- c()
for (year in 1990:2015){
  plantstemp <- plants %>% filter(year_built<=year & (year_retired>year | is.na(year_retired)==TRUE))
  plantstemp <- plantstemp %>%
                  group_by(plant_id) %>%
                  mutate(capacity = sum(capacity)) %>%
                  filter(row_number()==1) %>%
                  ungroup() %>%
                  dplyr::select(plant_id, capacity, lat, lon)
  plantstemp$year <- year
  plantsyearly <- rbind(plantsyearly, plantstemp)
}


# Tables
plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yield1reducedform <- readRDS("pollution_development/draft/tables/yield1reducedform.rds")
yield2naive <- readRDS("pollution_development/draft/tables/yield2naive.rds")
yield3ivmainfirststage <- readRDS("pollution_development/draft/tables/yield3ivmain_firststage.rds")
yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
yield3ivmain_lead <- readRDS("pollution_development/draft/tables/yield3ivmain_lead.rds")
yield4monthly <- readRDS("pollution_development/draft/tables/yield4monthly.rds")
yield5heterogeneity <- readRDS("pollution_development/draft/tables/yield5heterogeneity.rds")
yield6leads <- readRDS("pollution_development/draft/tables/yield6leads.rds")
yield7ivdiffindiff <- readRDS("pollution_development/draft/tables/yield7ivdiffindiff.rds")
yield8cumulative <- readRDS("pollution_development/draft/tables/yield8cumulative.rds")
labor1all <- readRDS("pollution_development/draft/tables/labor1all.rds")
labor2rural <- readRDS("pollution_development/draft/tables/labor2rural.rds")
labor3month <- readRDS("pollution_development/draft/tables/labor3month.rds")
laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
monthplot <- readRDS("pollution_development/draft/tables/monthplot.rds")
opplot <- readRDS("pollution_development/draft/tables/opplot.rds")
coef_vec <- readRDS("pollution_development/draft/tables/randomization_coef_vec.rds")
coefs_distfe <- readRDS("pollution_development/draft/tables/randomization_coefs_distfe.rds")
true_value <- readRDS("pollution_development/draft/tables/randomization_true_value.rds")
coef_current <- readRDS("pollution_development/draft/tables/coef_current.rds")
coef_lead <- readRDS("pollution_development/draft/tables/coef_lead.rds")
yieldrice <- readRDS("pollution_development/draft/tables/yieldrice.rds")
yieldricestatefe <- readRDS("pollution_development/draft/tables/yieldricestatefe.rds")


plants1990_capacity <- plants %>% filter(year_built<=1990 & (year_retired>1990 | is.na(year_retired)==TRUE))
plants2010_capacity <- plants %>% filter(year_built<=2010 & (year_retired>2010 | is.na(year_retired)==TRUE))

```


```{css, echo=F}
    /* Table width = 100% max-width */

    <!-- .remark-slide table{ -->
    <!--     width: 100%; -->
    <!-- } -->

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(n) {
      background-color: white;
    }
    
    tfoot {
      font-size: 80%;
    }

    table{
      border-collapse: collapse;
      border-color: transparent;
      background-color: "#FFFFFF";
    }
```


```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)

style_duo(
          primary_color = "#A7A9AC", 
          secondary_color = "#006334",
          table_row_even_background_color = "#FFFFFF"
          )

```





## This paper

- Effects of pollution on agricultural productivity in India
  - Previous evidence from gold mines (Aragon and Rud 2016) and workers in California (Graff Zivin and Neidell 2012)
- Higher exposure to pollution leads to lower agricultural productivity
  - Exposure defined as wind direction from coal plants towards villages
  - Also instrument for particulate matter 2.5 (PM 2.5)



---

## This paper

- Larger effects in areas growing more labor-intensive crops
  - Effects through labor?
- Compounding effects of shocks
  - Weather shocks + pollution = worse than either on its own





---

## Outline

<br><br>
- General idea and data<br><br>
- Validating wind direction<br><br>
- Estimating effects of pollution on agricultural productivity<br><br>
- Heterogeneity, including by labor intensities of crops<br><br>
- Overall pollution vs. pollution specifically from coal







---

## General idea
- Identify high-pollution locations based on where coal plants open or will open
  - Coal plants open in areas where pollution is already high
- Plot wind direction each day
  - Exposure means wind blowing from high-pollution locations to village (within 30km)












---


## Wind direction - first 100 days of 2010 in Guna district (MP)

```{r windbase, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = districts, color = "gray85", fill = "gray80") +
    geom_sf(data = villageswind %>% filter(date=="2010-01-01"),
            alpha = .5) +
    labs(title = "date: 2010-01-01") +
    theme(legend.position = "none")



```





---


## Wind direction - first 100 days of 2010 in Guna district (MP)

```{r wind, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = villageswind, 
            aes(fill = indicator),
            alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = date,
                      cumulative = FALSE
                      ) +
    labs(title = "date: {current_frame}") +
    theme(legend.position = "none")



```





---

## Empirical strategy

First stage

$$pm_{it} = \alpha_{i} + \gamma_{t} + \phi wind_{it} + \psi f(weather_{it}) + \varepsilon_{it}$$

Second stage

$$yield_{it} = \delta_{i} + \eta_{t} + \beta \hat{pm}_{it} + \theta f(weather_{it}) + \upsilon_{it}$$


- i indexes villages (districts for labor)
- t indexes time (where time differs based on the analysis)
- standard errors clustered at village throughout
- common use of stars (one indicates 0.10) throughout







---

## Where do coal plants open?


```{r plantresultstable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(plantresultstable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      plantresultstable,
      align = "cccc", linesep = "", escape = FALSE, format = "html"
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(10, hline_after = TRUE) %>%
  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
  kable_classic_2()

```







---

## Naive regression - yield on pollution


```{r naivetable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield2naive) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      yield2naive,
      align = "cccc", linesep = "", format = "html"
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2.5cm") %>%
  row_spec(6, bold = TRUE) %>%
  row_spec(8, hline_after = TRUE) %>%
  kable_classic_2()

```

- Likely endogenous
  - Upward biased?



---

## Reduced form - wind and yield


```{r yieldtable, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield1reducedform) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      yield1reducedform,
      align = "cccc", linesep = "", format = "html"
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2.5cm") %>%
  row_spec(6, bold = TRUE) %>%
  row_spec(8, hline_after = TRUE) %>%
  kable_classic_2()

```









---

## First stage - particulate matter and exposure


```{r yield3ivmainfirststage, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
yield3ivmainfirststage <- yield3ivmainfirststage[,-1]
# Table
colnames(yield3ivmainfirststage) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      yield3ivmainfirststage,
      align = "cccc", linesep = "", format = "html"
      ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(6, bold = TRUE) %>%
  row_spec(8, hline_after = TRUE) %>%
  kable_classic_2()

```










---

## Effects of PM on agricultural productivity

- We are really interested in the effect of _pollution_, not wind
- Use wind as instrument? Assumes...
  - Conditional on fixed effects and weather...
  - Wind only affects agriculture through pollution
- Reasonable?










---

## Effects of PM on agricultural productivity (IV)


```{r yieldtabletwo, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
# Table
yield3ivmain <- yield3ivmain[,-1]
colnames(yield3ivmain) <- c("(1)", "(2)", "(3)", "(4)")
rownames(yield3ivmain)[9] <- "F (first stage)"
kable(
      yield3ivmain,
      align = "cccc", linesep = "", format = "html"
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(6, bold = TRUE) %>%
  row_spec(8, hline_after = TRUE) %>%
  kable_classic_2()

```




---

## Robustness check: leads of wind direction


```{r leads, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
# Table
colnames(yield3ivmain_lead) <- c("(1)", "(2)")
rownames(yield3ivmain_lead)[9] <- "F (first stage)"
rownames(yield3ivmain_lead)[5] <- "weather (expanded, bins)"
yield3ivmain_lead[5,] <- c("Yes", "Yes")
kable(
      yield3ivmain_lead,
      align = "cc", linesep = "", format = "html"
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(6, bold = TRUE) %>%
  row_spec(8, hline_after = TRUE) %>%
  kable_classic_2()

```






---

## Heterogeneity

```{r heterogeneity, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield5heterogeneity) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
rownames(yield5heterogeneity)[c(1, 3)] <- c("PM 2.5", "PM times rain")
kable(
      yield5heterogeneity,
      align = "ccccc", linesep = "", format = "html"
      ) %>%
  column_spec(1, width = "4.5cm") %>%
  column_spec(c(2:6),width = "1.8cm") %>%
  row_spec(c(6), bold = TRUE) %>%
  row_spec(c(8), hline_after = TRUE) %>%
  add_header_above(c(" " = 1, "wind" = 2, "yield" = 2, " " = 1)) %>%
  kable_classic_2()

```





---

## Taking stock

- We see clear negative effects of pollution on productivity
  - Absolute deviation of exposure is approx. 8 days, meaning 2-3 percent change in productivity
  - Using AD of PM, it's above 20 percent change
- Key question: what is the driver?
  - Land?
  - Labor?
- Hard to look explicitly at land, so let's look at labor
  - Specifically, effects based on predominant crop


---


```{r crops, echo = FALSE, message = FALSE, warning = FALSE, dev='png', fig.height = 5, fig.cap = "Most common crop by village", fig.footer = list("Note: The figure shows the most common crop grown in each village, by area harvested.")}

village_shp <- village_shp %>% left_join(crops_all, by = c("shrid"))

# plot
ggplot() +
  geom_sf(data = india_shape, fill = "white") +
  geom_sf(data = village_shp, aes(fill = most_common), color = NA) +
  scale_fill_brewer("Most common crop\nby area harvested", palette = "Set3") +
  geom_sf(data = india_shape2, fill = NA, lwd = 0.01) +
  theme_minimal()

```



---

## How does this help?

- Different crops have different labor intensities
- Rice has the highest (from Michler, 2020):
  - Rice: 1,767 hours per hectare
  - Next highest: cotton at 857.5 hours per hectare







---

## Heterogeneity by rice growing regions

```{r yieldrice, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
yieldrice <- yieldrice[,-1]
colnames(yieldrice) <- c("(1)", "(2)")
kable(
      yieldrice,
      align = "cc", linesep = "", format = "html"
      ) %>%
  column_spec(1, width = "3.5cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(7, bold = TRUE) %>%
  row_spec(9, hline_after = TRUE) %>%
  kable_classic_2()

```




---

## Coal plants or overall pollution?

- I use the location of coal plants to identify high-pollution areas
- However, some coal plants open part-way through sample
- Differences-in-differences combined with IV
  - Differences in coal pollution, specifically, and other forms of pollution




---

## Coal plants over time

```{r plants, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = india_shape, color = "gray85", fill = "gray80") +
    geom_point(data = plantsyearly, 
               aes(x = lon, y = lat, size = capacity),
               colour = "#006334", alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = year,
                      cumulative = FALSE
                      ) +
    labs(title = "year: {current_frame}") +
    theme(legend.position = "none")



```







---

## Coal plants or overall pollution?


```{r yield7ivdiffindiff, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield7ivdiffindiff) <- c("(1)", "(2)")
kable(
      yield7ivdiffindiff,
      align = "cc", linesep = "", format = "html"
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(7, bold = TRUE) %>%
  row_spec(9, hline_after = TRUE) %>%
  kable_classic_2()

```





---

## Conclusion

- Pollution highest in poorer countries
  - South Asia, in particular
- Drivers of pollution are many
  - Coal plants
  - Transportation
  - Industry
  - Agriculture
  - etc.
- We already have plenty of evidence on effects of pollution on health






---

## Conclusion

- Paper presents evidence that air pollution leads to large decreases in agricultural productivity
  - Identification relies on within-village changes across seasons
- Air pollution worsened by around 15% from 1998 to 2019
  - Using most conservative result, this implies a decrease of around 9% in agricultural productivity
- Key finding: effects larger where crops are more labor intensive






---

## Conclusion

- Overall cost/benefit way beyond scope of paper
- Instead, results point to importance of the location of pollution sources
  - Downwind villages suffer
- 30km radius indicates relatively long-range effects






---

class: center, middle

<font size = "40"> Thank you! </font>

[https://joshmerfeld.github.io](https://joshmerfeld.github.io) <br>
Twitter: [@Josh\_Merfeld](twitter.com/Josh_Merfeld)








